---
title: "Integrating crew.ssh into a targets project"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Integrating crew.ssh into a targets project}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

This guide shows how to wire `crew.ssh` into a [`targets`](https://docs.ropensci.org/targets/)
project so that heavy targets run across a small SSH cluster. It assumes the
cluster itself is already reachable (passwordless SSH, shared/identical project
path on every node); see `vignette("setup")` for that. The patterns here suit an
[`renv`](https://rstudio.github.io/renv/) project that vendors its packages,
optionally as git submodules - a common layout - but adapt freely.

`crew.ssh` is **experimental**; expect breaking changes.

## The shape of the integration

A working setup has five pieces:

1. **`crew.ssh` installed** in the project library (and recorded in `renv.lock`).
2. **`_hosts.R`** - a gitignored, control-node-only file holding the cluster
   config as `crew.ssh.*` options. The single source of truth.
3. **`_hosts.R.example`** - a committed template.
4. **A node-provisioning helper** (`scripts/sync-nodes.R`) that pushes the project
   + packages to each node. A ready template ships with this package.
5. **`_targets.R` wiring** that reads `_hosts.R` and builds the controller.

Only `_hosts.R` is project-specific; everything else is reusable.

## 1. Install crew.ssh (and record it in renv.lock)

If you vendor packages from GitHub, install + record it in one step:

```{r}
renv::install("youruser/crew.ssh", lock = TRUE, prompt = FALSE)
```

`lock = TRUE` adds **only** crew.ssh to `renv.lock` (targeted). Then commit
`renv.lock`. The worker nodes will get crew.ssh when their `renv::restore()` runs
(step 4).

**Do not** run a plain `renv::snapshot()` to add it: in a project whose
lockfile contains development-only packages (testthat, roxygen2, devtools, ...
which show as `used = n` in `renv::status()`), the default implicit snapshot will
**prune** them from `renv.lock`. Use `renv::install(..., lock = TRUE)` or
`renv::record(...)` for targeted, additive lockfile edits, and keep
`git diff renv.lock` as a safety net.

(If crew.ssh is a local source tree rather than a GitHub repo, use
`renv::install("./path/to/crew.ssh")` - note the leading `./`.)

## 2. `_hosts.R` (the single source of truth)

Create `_hosts.R` at the project root and **gitignore it** (`echo _hosts.R >> .gitignore`).
It lives only on the control node; worker nodes never read it. All options use the
project-agnostic `crew.ssh.` prefix and are consumed by both `sync-nodes.R` and
your `_targets.R` wiring:

```{r}
# _hosts.R
options(
  crew.ssh.nodes   = c(node1 = 8L, node2 = 16L), # host = per-node worker capacity
  crew.ssh.projdir = "/home/you/myproject",       # identical path on every node
  crew.ssh.tunnel  = TRUE,                         # reverse-tunnel dial-back (firewall-free)
  crew.ssh.branch  = "main",                       # branch sync-nodes fast-forwards
  crew.ssh.symlinks = c(                           # symlinks sync-nodes (re)creates in projdir
    inputs  = "/mnt/shared/myproject/inputs",
    outputs = "/mnt/shared/myproject/outputs"
  )
  # direct dial-back only (crew.ssh.tunnel = FALSE):
  # crew.ssh.host = "control-node",  # control address reachable from workers
  # crew.ssh.port = 5555L,           # FIXED dial-back port (open it in the firewall)
  # optional: override the remote Rscript (default tracks the renv.lock R version):
  # crew.ssh.rscript = "Rscript-4.6.0"
)
```

Commit a sanitized `_hosts.R.example` (the block above with placeholder values) so
collaborators know the schema.

Per-node capacity (`caps`) is whatever a node can safely run at once; for RAM-bound
tasks, `cap = floor(usable_RAM / per_task_RAM)`. Because the dispatcher mixes tasks
across nodes and is not memory-aware, size a single controller for the heaviest
task, or split heavy and light work into separate controllers.

## 3. Provision the nodes (sync-nodes helper)

A node-provisioning helper ships with this package. Copy it into your project:

```{r}
file.copy(
  system.file("templates/sync-nodes.R", package = "crew.ssh"),
  "scripts/sync-nodes.R"
)
```

It reads `_hosts.R` and, for each node over SSH: fast-forwards `crew.ssh.branch`,
updates git submodules, (re)creates `crew.ssh.symlinks`, and runs
`renv::restore()` (which installs/updates whatever `renv.lock` specifies, including
crew.ssh once recorded). It is non-destructive (reports FAILED rather than
resetting), reads the R version from `renv.lock` (so an R upgrade is a one-line
change), and supports `--dry-run`. Run it from the project root on the control
node (after pushing, so the nodes can fetch):

```sh
Rscript scripts/sync-nodes.R --dry-run   # preview the remote script
Rscript scripts/sync-nodes.R             # apply
```

Each node needs git access to any private submodules, the project at
`crew.ssh.projdir`, and a per-node `renv` library (do not share an renv *library*
over a network filesystem across machines).

## 4. Wire `_targets.R`

Read `_hosts.R`, build the SSH controller when nodes are configured, and fall back
to a local controller otherwise (so a laptop / CI run still works). Keep light or
shared-service pools (downloads, etc.) local on the control node.

```{r}
if (file.exists("_hosts.R")) source("_hosts.R")

ssh_nodes <- getOption("crew.ssh.nodes")
controller <-
  if (is.null(ssh_nodes)) {
    crew::crew_controller_local(name = "primary", workers = 4L)
  } else {
    crew::crew_controller_group(
      crew.ssh::crew_controller_ssh(
        name    = "primary",
        nodes   = ssh_nodes,
        projdir = getOption("crew.ssh.projdir"),
        tunnel  = getOption("crew.ssh.tunnel", TRUE),
        host    = getOption("crew.ssh.host"),    # ignored when tunnel = TRUE
        port    = getOption("crew.ssh.port"),    # ignored when tunnel = TRUE
        rscript = getOption("crew.ssh.rscript"), # NULL -> the session's R (homogeneous)
        seconds_idle = Inf,                      # persistent workers
        crashes_max  = 25L
      ),
      crew::crew_controller_local(name = "download", workers = 8L, seconds_idle = 60)
    )
  }

targets::tar_option_set(controller = controller)
```

Route a target to a pool with
`tar_resources(crew = tar_resources_crew(controller = "primary"))`. Note that
`tar_resources_crew(controller =)` is length-1: a branched target goes to ONE
controller, so spreading one heavy branched target across nodes is done by the
single capacity-weighted `primary` controller, not by per-node controllers.

## 5. Verify

Preflight (read-only) before any real run:

```{r}
crew.ssh::crew_ssh_check(
  nodes   = getOption("crew.ssh.nodes"),
  projdir = getOption("crew.ssh.projdir"),
  packages = c("crew", "targets")
)
#> crew.ssh preflight: 2/2 node(s) OK
```

Then a small smoke test of the full path:

```{r}
ctrl <- crew.ssh::crew_controller_ssh(
  nodes = getOption("crew.ssh.nodes"),
  projdir = getOption("crew.ssh.projdir"),
  tunnel = TRUE, seconds_idle = 10
)
ctrl$start()
for (i in 1:6) ctrl$push(command = Sys.info()[["nodename"]], name = sprintf("t%d", i))
ctrl$wait(seconds_timeout = 120)
res <- character(0)
repeat {
  r <- ctrl$pop()
  if (is.null(r)) break
  res <- c(res, r$result[[1]])
}
ctrl$terminate()
length(res) # expect 6, spread across > 1 node
```

## Notes and gotchas

- **Control node vs workers.** The control node runs `tar_make` / the dispatcher
  and needs `crew.ssh`; it does not need to run sshd (it dials out). Worker nodes
  need `crew` plus every package your tasks use; they do not strictly need
  `crew.ssh`, though `renv::restore()` installs it everywhere if it is in the lock.
- **Dial-back is not SSH.** A worker's connection back to the dispatcher is a raw
  TCP socket, so it ignores `~/.ssh/config` / custom SSH ports. On a firewalled
  network use `tunnel = TRUE` (no inbound port) rather than opening ports; see
  `vignette("setup")`.
- **Local packages from GitHub.** If your vendored packages are recorded in
  `renv.lock` as `Source: GitHub`, `renv::restore()` installs them from GitHub on
  each node; `renv::install("path")` for a git-submodule path can fail (renv tries
  to resolve it as a remote) - use the `"user/repo"` spec.
- **Re-provision after package changes.** When a vendored package changes, record
  the new version (`renv::install("user/pkg", lock = TRUE)`), commit + push
  `renv.lock`, then re-run `sync-nodes.R` so the nodes pick it up.
- See also `vignette("setup")` for SSH keys, `known_hosts`, `~/.ssh/config`, the
  reverse tunnel, and the env-driven integration tests.
```
