---
title: "Setting up an SSH cluster for crew.ssh"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting up an SSH cluster for crew.ssh}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

`crew.ssh` is **experimental**.
It launches each worker as a remote R process over SSH and assumes a **trusted local network**.
It does no key management or brokering of its own: it shells out to your system `ssh` client, which uses your existing keys, agent, and `~/.ssh/config`.
This guide covers getting a small cluster ready.
The examples use `host1`, `host2`, etc. as placeholder machine names.

## What every node needs

1. **Passwordless SSH** from the control machine to each node (keys, below).
2. **The same project and R library** reachable at the same `projdir` path on every node, for example a shared network filesystem (NFS) or identical local checkouts with identical package versions.
   A worker runs `cd <projdir>` first, so the project `.Rprofile` / `renv` activates and the worker finds the packages it needs.
3. **A reachable dial-back address.**
   Workers connect back to the controller, so the `host` you pass must be reachable *from the worker nodes* (not `127.0.0.1`).

## 1. Create an SSH key (if you do not have one)

On the control machine:

```bash
ssh-keygen -t ed25519 -C "you@example.org"
# press Enter to accept the default path (~/.ssh/id_ed25519)
# set a passphrase, then load it once per session with ssh-agent (below)
```

Use an agent so the passphrase is entered once, not per worker launch:

```bash
eval "$(ssh-agent -s)"
ssh-add ~/.ssh/id_ed25519
```

## 2. Authorize the key on each node

```bash
ssh-copy-id you@host1
ssh-copy-id you@host2
# ... one per node
```

If the nodes share your home directory over NFS, a single `ssh-copy-id` (or appending the public key to `~/.ssh/authorized_keys` once) covers them all.

## 3. Record host keys (known_hosts)

`crew.ssh` passes `-o BatchMode=yes` by default, so SSH never prompts.
That also means an **unknown host key makes the connection fail** rather than asking you to accept it.
Record each node's host key first:

```bash
ssh-keyscan -H host1 host2 >> ~/.ssh/known_hosts
```

Keep host-key checking on (do not set `StrictHostKeyChecking=no`): it is your protection against a machine-in-the-middle on the LAN.

## 4. Use names, not just IPs

The per-node `host` is handed straight to `ssh`, so anything `ssh` can resolve works: a DNS name, an `/etc/hosts` entry, or - best - an `~/.ssh/config` alias.
Aliases let you pin the user, key, and port per host without changing your R code:

```
# ~/.ssh/config
Host host1
    HostName 192.168.1.21
    User you
    IdentityFile ~/.ssh/id_ed25519
```

With that, `crew_ssh_node("host1", ...)` just works.

One caveat for the **dial-back `host`** (the control-node address workers connect back to): it must resolve **on the worker nodes**, not only on the control machine.
A name that only exists in the control node's `/etc/hosts` will not help the workers.
Use a name every node resolves consistently (shared DNS or matching `/etc/hosts` everywhere), or the control node's IP address.

## 5. Verify before you run

A quick manual check:

```bash
ssh -o BatchMode=yes host1 true && echo OK
```

Then use the built-in preflight, which checks connectivity, the project directory, the resolved `Rscript`, and that the required packages load on every node at once:

```{r}
library(crew.ssh)

crew_ssh_check(
  nodes = c(host1 = 1L, host2 = 1L),
  projdir = "/home/you/myproject",
  packages = c("crew", "targets")
)
#> crew.ssh preflight: 2/2 node(s) OK
#>    host  ssh projdir rscript r_version missing_packages   ok detail
#> 1 host1 TRUE    TRUE    TRUE     4.6.0                  TRUE
#> 2 host2 TRUE    TRUE    TRUE     4.6.0                  TRUE
```

If a node fails, the `detail` column carries the first error (for example a missing project directory, or `Rscript` not found on a heterogeneous node - set `rscript` for that node via `crew_ssh_node()`).

## 6. The R binary on each node

By default (`homogeneous = TRUE`) `crew.ssh` runs the **launching session's own** `Rscript` (`file.path(R.home("bin"), "Rscript")`) on every node, mirroring `parallelly::makeClusterPSOCK()`.
This is a full path, so it does not depend on the remote `PATH` or on any version-suffixed name.
It is correct when every node has the same R installation path.

If a node's R lives elsewhere, override it for that node:

```{r}
crew_ssh_node("host3", workers = 8L, rscript = "/opt/R/4.6.0/bin/Rscript")
```

Or set `homogeneous = FALSE` to use bare `Rscript` from each node's `PATH` (note that non-interactive SSH sessions often have a reduced `PATH`).

## 7. Put it together

```{r}
library(crew.ssh)

controller <- crew_controller_ssh(
  name = "workers",
  nodes = c(host1 = 88L, host2 = 95L),
  projdir = "/home/you/myproject", # same path on every node
  host = "control-node" # reachable from the worker nodes
)

# targets:
# targets::tar_option_set(controller = controller)

# or drive crew directly:
controller$start()
controller$push(command = Sys.info()[["nodename"]], name = "where")
controller$wait()
controller$pop()$result[[1]] # the worker node's hostname
controller$terminate()
```

## Worker teardown and cleanup

On a normal stop the worker exits gracefully, and its `seconds_idle` /
`seconds_wall` timeouts bound how long a stranded worker can survive.
A *hard* kill of the local `ssh` client does not by itself guarantee the remote R exits.
For prompt teardown in that case, launch with a forced pseudo-tty:

```{r}
controller <- crew_controller_ssh(
  nodes = c(host1 = 88L, host2 = 95L),
  projdir = "/home/you/myproject",
  host = "control-node",
  request_tty = TRUE # ssh -tt: remote R gets SIGHUP when the ssh client dies
)
```

The trade-off is that `-tt` merges and alters worker stdout-stderr capture, so it is off by default.

One boundary to keep in mind: neither SSH nor `crew` reaps **child processes the worker spawns**, such as a `docker run` container.
If a task launches one, the task code must tear it down itself (for example `docker run --rm` plus stopping the container when the R worker exits).

## Firewalled networks: reverse-tunnel dial-back

By default a worker dials the dispatcher back over a *direct* TCP connection
(`tcp://<host>:<port>`).
That is not SSH, so it ignores `~/.ssh/config` and any custom SSH ports, and it
needs that port reachable from the worker nodes - a problem behind a firewall
(`ufw` / `firewalld`).

Set `tunnel = TRUE` to dial back through an **SSH reverse tunnel** instead:

```{r}
crew_controller_ssh(
  nodes = c(host1 = 8L, host2 = 16L),
  projdir = "/home/you/myproject",
  tunnel = TRUE
)
```

The dispatcher then binds `127.0.0.1`, and each worker is launched with
`ssh -R <node-port>:127.0.0.1:<dispatcher-port>`, so the dial-back rides the
existing SSH connection.
No inbound firewall port is opened, custom SSH ports keep working (the launch is
still a plain `ssh <host>`), and the dispatcher is not exposed on the network at
all.
This is the recommended mode on locked-down clusters.

## Security notes

- **Keys are never handled by the package.**
  It uses your system `ssh` (keys in `~/.ssh`, `ssh-agent`, `~/.ssh/config`).
  Protect keys as usual: a passphrase plus an agent, correct file permissions, and per-user keys.
- **`BatchMode=yes`** (a default ssh option) means no password prompts and fail-fast behaviour; combined with a populated `known_hosts` it also prevents silent host-key changes.
- **The dial-back channel** carries serialized R objects.
  On anything less than a fully trusted network, encrypt it by passing `tls = crew::crew_tls(mode = "automatic")` to `crew_controller_ssh()`.
- **Command-line exposure.**
  The worker launch command (which includes the dispatcher URL and any TLS token) is briefly visible in the remote process list (`ps`).
  This is not your private key, but is worth knowing on a shared node.

## Testing on a real cluster

The package's SSH tests are dormant by default (so `R CMD check` and CI never touch a cluster) and are configured entirely through the environment - no hostnames are hard-coded.

A gated multi-node test (`tests/testthat/test-integration-cluster.R`) confirms that work actually spreads across more than one node.
Enable it by setting:

```sh
export CREW_SSH_TEST_NODES="host1=2,host2=2"   # host = per-node cap
export CREW_SSH_TEST_PROJDIR="/path/to/project" # shared/identical path on the nodes
export CREW_SSH_TEST_HOST="control-node"        # reachable from the worker nodes
# optional: export CREW_SSH_TEST_RSCRIPT="/opt/R/4.6.0/bin/Rscript"
```

then run the test suite on the control node.

For the fault-injection / orphan exercise (kill a worker mid-run, then check for lingering remote processes), use the manual soak script, which reads the same variables:

```sh
Rscript -e 'source(system.file("soak/run_soak.R", package = "crew.ssh"))'
```

See `inst/soak/README.md` for what to look for.
The single-machine loopback test (`tests/testthat/test-integration-loopback.R`) is separate and enabled with `CREW_SSH_TEST_INTEGRATION` plus working passwordless `ssh localhost`.
